---
title: "Supermarket Sales Analysis"
author: "Vaghasiya Meet R."
output: html_document
---

# Introduction

This project analyzes supermarket sales data using R.

The main purpose of this project is to understand sales,
product performance, customer information, and payment methods.

# Objectives

- To import supermarket sales data.
- To clean and preprocess the data.
- To perform statistical analysis.
- To create graphs and plots.
- To interpret the results.
- To identify important findings from the data.

# Data Importing

First, we import the supermarket sales dataset into R.

```{r}
data <- read.csv("supermarket_sales.csv.csv")

head(data)
```

# Data Cleaning and Preprocessing

First, we check the structure of the dataset.

```{r}
str(data)
```

Now, we check for missing values.

```{r}
colSums(is.na(data))
```

Now, we check for duplicate records.

```{r}
sum(duplicated(data))
```

Finally, we display the summary of the dataset.

```{r}
summary(data)
```

# Data Analysis

## Total Sales

```{r}
total_sales <- sum(data$Total_Sales)
total_sales
```

## Average Sales

```{r}
average_sales <- mean(data$Total_Sales)
average_sales
```

## Maximum Sales

```{r}
maximum_sales <- max(data$Total_Sales)
maximum_sales
```

## Minimum Sales

```{r}
minimum_sales <- min(data$Total_Sales)
minimum_sales
```

## Total Quantity Sold

```{r}
total_quantity <- sum(data$Quantity)
total_quantity
```

# Graphs and Plots

## 1. Sales by Product Line

```{r}
sales_by_product <- aggregate(Total_Sales ~ Product_Line, data = data, sum)

barplot(
  sales_by_product$Total_Sales,
  names.arg = sales_by_product$Product_Line,
  main = "Total Sales by Product Line",
  xlab = "Product Line",
  ylab = "Total Sales"
)
```

## 2. Sales by Branch

```{r}
sales_by_branch <- aggregate(Total_Sales ~ Branch, data = data, sum)

barplot(
  sales_by_branch$Total_Sales,
  names.arg = sales_by_branch$Branch,
  main = "Total Sales by Branch",
  xlab = "Branch",
  ylab = "Total Sales"
)
```

## 3. Sales Distribution

```{r}
hist(
  data$Total_Sales,
  main = "Distribution of Total Sales",
  xlab = "Total Sales",
  ylab = "Frequency"
)
```

## 4. Quantity vs Total Sales

```{r}
plot(
  data$Quantity,
  data$Total_Sales,
  main = "Quantity vs Total Sales",
  xlab = "Quantity",
  ylab = "Total Sales"
)
```

## 5. Sales by Payment Method

```{r}
sales_by_payment <- aggregate(Total_Sales ~ Payment, data = data, sum)

barplot(
  sales_by_payment$Total_Sales,
  names.arg = sales_by_payment$Payment,
  main = "Total Sales by Payment Method",
  xlab = "Payment Method",
  ylab = "Total Sales"
)
```
# Interpretation of Results

## Product Line Analysis

The bar chart shows the total sales for different product lines.
It helps us compare which product categories generate more sales.

## Branch Analysis

The branch-wise sales chart shows the difference in sales between
different supermarket branches.

## Sales Distribution

The histogram shows how the total sales values are distributed.
It helps us understand the general pattern of sales.

## Quantity and Sales Analysis

The scatter plot shows the relationship between quantity sold and
total sales. It helps us understand whether higher quantities are
associated with higher sales.

## Payment Method Analysis

The payment method chart shows the total sales made through different
payment methods such as Cash, Card, and UPI.
# Findings

The analysis of supermarket sales data provides the following findings:

- Different product lines have different levels of sales.
- Sales vary among the different supermarket branches.
- The quantity of products sold has a relationship with total sales.
- Customers use different payment methods such as Cash, Card, and UPI.
- Statistical analysis helps us understand the overall sales performance.
- Graphs make it easier to compare sales and identify patterns in the data.

# Conclusion

This project analyzed supermarket sales data using R.

The data was imported, cleaned, and statistically analyzed.
Different graphs were created to understand sales patterns,
product performance, branches, quantity, and payment methods.

R is useful for analyzing data and presenting the results
in a simple and understandable form.